### 2019-10-01 paper.tex
first commit

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage{array,multirow}
\usepackage{balance}
% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}
%\usepackage{setspace}
\pagestyle{empty} 
%\doublespacing

% DOI
%\acmDOI{10.475/123_4}

% ISBN

\begin{document}
\title{A Cross-Domain Exploratory Analysis of \\ Social Media Data During Crisis Events}

\author{Hernan Sarmiento}
\affiliation{%
  \institution{Department of Computer Science \\ University of Chile}
  %\streetaddress{P.O. Box 1212}
   \city{Santiago} 
   \country{Chile}}
\email{hsarmien@dcc.uchile.cl}

\author{Barbara Poblete}
\affiliation{%
  \institution{Department of Computer Science \\ University of Chile}
  %\streetaddress{P.O. Box 1212}
   \city{Santiago} 
   \country{Chile}}
\email{bpoblete@dcc.uchile.cl}

\begin{abstract}

\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10002951.10003227.10003351.10003446</concept_id>
	<concept_desc>Information systems~Data stream mining</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10002951.10003227.10003236</concept_id>
	<concept_desc>Information systems~Spatial-temporal systems</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Data stream mining}
\ccsdesc[300]{Information systems~Spatial-temporal systems}

\keywords{Emergency Situations; Citizen Sensors; Social Media}

\maketitle

\section{Introduction}

\section{Data Acquisition}

\begin{itemize}
	\item CrisisLex and CrisisNlp.
	\item XX dataset with YY instances and ZZ events.
\end{itemize}

\section{Data Preparation}

\begin{itemize}
	\item Remove duplicates
	\item Remove data without metadata
\end{itemize}

\subsection{Detecting Languages}

\subsection{Adding country information}

\subsubsection{Adding primary languages}

\subsection{Merging Labels}

\subsection{Adding Crisis Categories}

\section{Exploratory Analysis}

\subsection{Computing Features}

\subsubsection{Content Features}

\subsubsection{Grammatical Features}

\subsubsection{Entities-Based Features}

\subsubsection{Emotional Features}

\subsection{Hazard Categories Analysis}

\subsection{Hazard Sub Categories Analysis}

\subsection{Hazard Types Analysis}

\begin{itemize}
	\item $82\%$ english data and $8\%$ spanish and italian each.
	\item Most events ocurred in United States
\end{itemize}

\begin{acks}
This work has been partially funded by the Millennium Institute for Foundational Research on Data.	
\end{acks}

\bibliographystyle{ACM-Reference-Format}
\balance
\bibliography{bio}

\end{document}